# Support Vector Machine (SVM)

## R Packages and Setup {-}

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
library(kernlab)
```


## Introduction

Below paragraph is copied from the assignment, we are planning on updating this as we formulate our narrative for the final deliverable:

Support Vector Machines (SVMs) are algorithms for classification and regression tasks. Linear (Vanilla Dot) and Radial Basis Function (RBF) kernels are popular choices. The Vanilla Dot kernel is simple, effective for linearly separable data, but can’t handle complex non-linear relationships. RBF transforms data into a high-dimensional space to model non-linear relationships but requires careful tuning for optimal performance.

In SVM, we will use both Vanilla SVM and RBF SVM. Our objective is to evaluate the performance of these two methods in terms of accuracy and sensitivity when predicting our test dataset.


## Load Data

```{r}
# Load data
train_svm <- read.csv('Train Test Set/train_svm.csv')
test_svm <- read.csv('Train Test Set/test_svm.csv')

# Display statistics
str(train_svm)
summary(train_svm)
head(train_svm)
```


## Model SVM

### Vanilla SVM

```{r model_svm_vanilla, cache=TRUE, message=FALSE, warning=FALSE, results='hide'}
# Build a vanilla model
model_svm_vanilla <- ksvm(obesity_leveloverweight ~ .,
                          data = train_svm,
                          kernel = 'vanilladot')
```

```{r model_svm_vanilla_summary}
# Display summary
summary(model_svm_vanilla)
```

### RBF SVM

```{r model_svm_rbf, cache=TRUE, message=FALSE, warning=FALSE, results='hide'}
# Build a rbf model
model_svm_rbf <- ksvm(obesity_leveloverweight ~ .,
                      data = train_svm,
                      kernel = 'rbfdot')
```

```{r model_svm_rbf_summary}
# Display summary
summary(model_svm_rbf)
```


## Evaluate SVM

### Vanilla SVM

```{r evaluate_svm_vanilla}
# Make a prediction
prediction_svm_vanilla_probability <- predict(model_svm_vanilla, test_svm)
prediction_svm_vanilla <- ifelse(prediction_svm_vanilla_probability >= 0.5, 1, 0)
summary(prediction_svm_vanilla_probability)

# Perform confusion matrix
cm_svm_vanilla <- confusionMatrix(as.factor(prediction_svm_vanilla),
                                  as.factor(test_svm$obesity_leveloverweight),
                                  positive = '1')
cm_svm_vanilla
```

Based on the confusion matrix, we observe that the model achieves an accuracy of ``r round(cm_svm_vanilla$overall[1], 4)``, sensitivity of ``r round(cm_svm_vanilla$byClass[1], 4)``, and a Kappa coefficient of ``r round(cm_svm_vanilla$overall[2], 4)``. We will assess these results towards the conclusion of the assignment when we have completed stacked models.


```{r}
pred <- ROCR::prediction(prediction_svm_vanilla_probability, test_svm$obesity_leveloverweight)
perf <- ROCR::performance(pred, measure = "tpr", x.measure = "fpr")
auc <- ROCR::performance(pred, measure="auc")


auc <- round(auc@y.values[[1]],3)
roc.data <- data.frame(fpr=unlist(perf@x.values),
                       tpr=unlist(perf@y.values),
                       model="GLM")

ggplot(roc.data, aes(x=fpr, ymin=0, ymax=tpr)) +
    geom_ribbon(alpha=0.2, fill = "blue") +
    geom_line(aes(y=tpr), col = "blue") +
    geom_abline(intercept = 0, slope = 1, lty = "dashed") +
    labs(title = paste0("Vanilla SVM ROC Curve w/ AUC=", auc)) +
    theme_bw()
```

```{r write_prediction_svm_vanilla, echo=FALSE}
# Write to csv
write.csv(prediction_svm_vanilla,
          file = 'Predictions/prediction_svm_vanilla.csv',
          row.names = FALSE)

# Extract confusion matrix
write.csv(cm_svm_vanilla$table,
          file = 'Confusion Matrix/cm_svm_vanilla.csv',
          row.names = FALSE)
```

### RBF SVM

```{r evaluate_svm_rbf}
# Make a prediction
prediction_svm_rbf_probability <- predict(model_svm_rbf, test_svm)
prediction_svm_rbf <- ifelse(prediction_svm_rbf_probability >= 0.5, 1, 0)

# Perform confusion matrix
cm_svm_rbf <- confusionMatrix(as.factor(prediction_svm_rbf),
                              as.factor(test_svm$obesity_leveloverweight),
                              positive = '1')
cm_svm_rbf
```

Based on the confusion matrix, we observe that the model achieves an accuracy of ``r round(cm_svm_rbf$overall[1], 4)``, sensitivity of ``r round(cm_svm_rbf$byClass[1], 4)``, and a Kappa coefficient of ``r round(cm_svm_rbf$overall[2], 4)``. We will assess these results towards the conclusion of the assignment when we have completed stacked models.

```{r}
pred <- ROCR::prediction(prediction_svm_rbf_probability, test_svm$obesity_leveloverweight)
perf <- ROCR::performance(pred, measure = "tpr", x.measure = "fpr")
auc <- ROCR::performance(pred, measure="auc")

auc <- round(auc@y.values[[1]],3)
roc.data <- data.frame(fpr=unlist(perf@x.values),
                       tpr=unlist(perf@y.values),
                       model="GLM")

ggplot(roc.data, aes(x=fpr, ymin=0, ymax=tpr)) +
    geom_ribbon(alpha=0.2, fill = "blue") +
    geom_line(aes(y=tpr), col = "blue") +
    geom_abline(intercept = 0, slope = 1, lty = "dashed") +
    labs(title = paste0("RBF SVM ROC Curve w/ AUC=", auc)) +
    theme_bw()
```

```{r write_prediction_svm_rbf, echo=FALSE}
# Write to csv
write.csv(prediction_svm_rbf,
          file = 'Predictions/prediction_svm_rbf.csv',
          row.names = FALSE)

# Extract confusion matrix
write.csv(cm_svm_rbf$table,
          file = 'Confusion Matrix/cm_svm_rbf.csv',
          row.names = FALSE)
```

